---
title: "Lab Report Ver2"
author: "Aaron Willcox"
date: "03/08/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Included Library, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Included librarys for functions
library(psych)
library(haven)
library(dplyr)
library(lmSupport)
library(stargazer)
library(tableone)
```

**DV:** Personal Wellbeing Index (pwi)  
**IVs:** Affect(affect HPMood), Cogniton(cognition, MDT), Gender(gender), Personality(personality)

***

```{r Import}
well_being_df <- read_sav("WellbeingFINAL.sav")

```

```{r Check Data}
nrows <- nrow(well_being_df)
ncomplete <- sum(complete.cases(well_being_df))
ncomplete/nrows

# Check for missing or na data
which(is.na(well_being_df))
```


As per the PWI manual: score/number of domains*10 to get a SWB score. 
Then mutate this to a new column called "PW.Index"
```{r PWI Score}
well_being_df <- well_being_df %>% 
  mutate(PW.Index = (
    PWI1 + 
      PWI2 + 
      PWI3 + 
      PWI4 + 
      PWI5 + 
      PWI6 + 
      PWI7)/7*10)
```


Personality scores for each trait(O,C,E,A,N) are taken for each group of 10. 
These are summed and mutated to a new column for each trait.

```{r Personality}
# Personality Traits
well_being_df <- well_being_df %>% 
  mutate(Openness = ( # Openness
    Personality1 + 
      Personality2 +
      Personality3 +
      Personality4 +
      Personality5 +
      Personality6 +
      Personality7 +
      Personality8 +
      Personality9 +
      Personality10), 
    Conscientiousness = ( # Conscientiousness
      Personality11 + 
        Personality12 +
        Personality13 +
        Personality14 +
        Personality15 +
        Personality16 +
        Personality17 +
        Personality18 +
        Personality19 +
        Personality20),
    Extraversion = ( # Extraversion
      Personality21 + 
        Personality22 +
        Personality23 +
        Personality24 +
        Personality25 +
        Personality26 +
        Personality27 +
        Personality28 +
        Personality29 +
        Personality30),
    Agreeableness = ( # Agreeableness
      Personality31 + 
        Personality32 +
        Personality33 +
        Personality34 +
        Personality35 +
        Personality36 +
        Personality37 +
        Personality38 +
        Personality39 +
        Personality40),
    Neuroticism  = (  # Neuroticism
      Personality41 + 
        Personality42 +
        Personality43 +
        Personality44 +
        Personality45 +
        Personality46 +
        Personality47 +
        Personality48 +
        Personality49 +
        Personality50))

```

Homeostatically Protected Mood (Core Affect) as a measuer by Davern., et, al. that asks how people 
feel about life in general across three domains of "happy", "content" & "excited/alert".
Here the relevant columns are pulled, summed, divided by the 3 domains and mutated to a new column
called "affect". 
```{r Affect}
well_being_df <- well_being_df %>% 
  mutate(affect = 
           (Affect1 + # happy
              Affect2 + # content
              Affect3)) # excited

```


Multiple Discrepancies Theory (MDT) - Cognition 
Firstly I mutate to a new column labelled "MDT" and averaged. 
Then each discrepancy mean is worked out and stored in its variable. 

1. Self Best
2. Self Future
3. Self Progress
4. Self Needs
5. Self Deserves
6. Self Other
7. Self Wants

```{r MDT}
# This will create a new variable with the overal MDT cognition mean score
well_being_df <- well_being_df %>%  
  mutate(MDT =  (
    Cognition1 +
      Cognition2 +
      Cognition3 +
      Cognition4 +
      Cognition5 +
      Cognition6 +
      Cognition7)/7
  )

# Rename columns to refelct the discrepancies 
well_being_df <- well_being_df %>% 
  mutate(self_best =     Cognition1,
         self_future =   Cognition2,
         self_progress = Cognition3,
         self_needs =    Cognition4,
         self_deserves = Cognition5,
         self_other =    Cognition6,
         self_wants =    Cognition7)



```

### Descriptives of key variables
```{r Descriptives}
Descriptives <- describe(well_being_df[71:85]) #Using the brackets to call on the rows that were mutated to the dataframe

# Making use of the stargazer library to output a nicer looking table for html
# Cummins et.al, 2009 showed that the population SWB mean 74.93 and SD = 12.36.
stargazer(Descriptives, 
          title = "Descriptives",
          type = "text", 
          style = "all", 
          summary = FALSE, 
          ci = FALSE, digits = 2)
```



Looking at the discriptives things look pretty ok in terms of skew.
The main response variable is negatively skewed however, given the population mean
is up around 70s then this looks ok. 
```{r Plots}
hist(well_being_df$PW.Index)
```



Next I will have a look at the distribution of the Big 5 scores
Graphical representations of Big 5 scores in a box plot. 

```{r Big 5 Plot}
bplot.big5 <- cbind(Openness, 
                    Conscientiousness, 
                    Extraversion, 
                    Agreeableness, 
                    Neuroticism)

boxplot.names <- c("O", "C", "E", "A", "N")

boxplot(bplot.big5, main = "Big 5 Means", names = boxplot.names, horizontal = FALSE)  

```


Combine into an object variables of interest. Using cbind as calling variables with well_being_df$... removes col names for some reason

```{r Correlations}
Group.Corr <- well_being_df[71:85] 

# Check all the correlations in a matrix
round(cor(Group.Corr, use = "pair"), 2) 

```


Openess was negatively assocaited with PWI scores (-0.32), neuroticism
had a slight negative association but the remainder of the big 5 traits had little association.
Which is surprising considering that if you are open trait then woulnt you have thought the SWB be higher?
Strongest correlation was that of affect followed by MDT. 
Break down of the discrepancies: Self.other had little or no association
on SWB. 

### Regression
Step one of the multiple regression will use the two strongest correlates followed by the big
5 then MDT. 

```{r Regression}
model1 <- lm(PW.Index ~ Age + Gender, data = well_being_df)

model2 <- lm(PW.Index ~ Age + Gender +
               affect + MDT, data = well_being_df) #Both sig effect of affect and MDT

model3 <- lm(PW.Index ~ Age + Gender +
               affect + MDT +
               Openness + Conscientiousness + 
               Extraversion + Agreeableness +
               Neuroticism, data = well_being_df) #additional of personality constructs - not sig

model4 <- lm(PW.Index ~ Age + #model without overall MDT effect and with discrepancies
               Gender +
               affect +
               self_best +
               self_future + 
               self_progress +
               self_needs + 
               self_deserves +
               self_other + 
               self_wants, data = well_being_df) 


model5 <- lm(PW.Index ~ Age + #model without overall MDT effect and with discrepancies and Big5
               Gender +
               affect +
               self_best +
               Openness + 
               Conscientiousness + 
               Extraversion + 
               Agreeableness +
               Neuroticism +
               self_future + 
               self_progress +
               self_needs + 
               self_deserves +
               self_other + 
               self_wants, data = well_being_df) #62% variance. self_best/future are sig

#removing Big Five constructs has 0.6066 then adding has 0.6169. 2% difference
#using only the extraveriosn and neurotic traits is where the 2% additoinal variance is coming form. 
#


#Hierachial regression 
summary(model1) #PWI ~ age and gender. No sig predictors r.sqrd = 0.016
summary(model2) #affect p<0.00001 and MDT p<0.01 r.sqrd = 0.55 
summary(model3) #Big 5 same sig as model2 r.sqrd = 0.5639 - ANOVA showed NO sig effect on model
summary(model4) #age p<.05, affect p<0.00001, self_best p<0.05, self_future p<0.01
summary(model5) #non sig
# MOdel for was the only sig model at p<0.01

# Run an ANOVA to check for sig effect. 
anova(model1, model2, model3, model4, model5)

effect_sizes_model4 <- modelEffectSizes(model4, Digits = 2) #check effect sizes and partials

# Print a nice table to screen using the stargazer library
stargazer(model4,
          title = "Results",
          type = "text", 
          style = "all", 
          summary = TRUE, 
          dep.var.labels = "Personal Wellbeing Index", 
          ci = FALSE, digits = 2) #the text prints table to the Console

```


Affect was the strongest predictor on SWB followed by self.future. 
Model 3 was the only significant model accounting for 62% variance in SWB and
affect was the strongest unique contributor to SWB followed by self.future. 


