---
title             : "Competing Models of SWB"
shorttitle        : "Models of SWB"

author: 
  - name          : "Aaron L. Willcox"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Deakin University"
  - id            : "2"
    institution   : "Konstanz Business School"



abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
source("analysis.R")
```
# Introduction
Subjective Wellbeing is the overall evaluation of feeling satisfied absed on cognitive, emotional domains. However, the contributions as to how much of each domain is still under much debate. The Australian Unity Wellbeing Index provides a snapshot of the overwellbeing of Australians each year. This study has a relateively robust score and has deomstrated that, on average, the population PWI average is 75 with a 3.0 percentage points difference. What this suggests is that a persons wellbeing has a set-point as to where their sense of happiness is and therefore can be used to evaluate wellbeing.  


# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

Measures of Employment history were also recorded but were not included in the analyis. 

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results
I need to print and omega symbol $\omega$ and I hope that this worked
maybe if i try this one  or perhaps this one 

```{r, results='asis'}
apa_table(H_regression$table, added_stub_head = "Variables", note = "This is a note section", font_size = "small")
```

Then on the next page we have here a loverly correlation table for me to check it and see if the covaiates match with the regression model
```{r, results='asis'}
apa_table(a
          , escape  = FALSE
          , format = "html"
          , caption = "Correlation matrix of the main variables"
          , note    = "p<.001***, p<.01**, p<.05*, p<.10+")
```


# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
