---
title             : "Competing Models of SWB"
shorttitle        : "Models of SWB"

author: 
  - name          : "Aaron L. Willcox"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"

affiliation:
  - id            : "1"
    institution   : "Deakin University"




abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
source("analysis.R")
```
# Introduction
Subjective wellbeing (SWB) is the term applied to explain an individual's experience and perceptions of how they evaluate their life across various domains. It is subjective in that it captures the individuals cognitive and affective components in how they think and feel in areas such as health, work and relationships. How these components relate and merge to form an overall measure of wellbeing can help determine what aspects of an individual’s lives are driving how they feel. @RN299 argues that SWB is predominately driven by a cognitive component and there is no single overarching model of wellbeing. However, @RN421 proposed a neurophysiological model of SWB analogous to how body temperature is regulated. The regulation of wellbeing should fall within a normative range of 73.43 - 76.43 points and values outside of this range may indicate problems. This suggests that the subjectiveness nature of wellbeing is innate and generally positive. In contrast to {Diener, 2018 #467}, external factors only play a small role in happiness {Lucas, 2008 #549}. Furthermore, the individual can adapt to challenging circumstances and return to a state of normal functioning [@RN383]. However, the extent to what components drive the model and which is the strongest, is still under debate.

Current theories suggest that affective factors, such as a person's mood and emotions, is the dominant component in the model {Davern, 2007 #417}. SWB is generally positive and maintained by an inherent set-point that is managed through an evolutionary system called Homeostatically Protected Mood (HPMood). When this system is disrupted through challenging experiences, the set-point is expected to drop below the average and experience negative affects rather than positive ones {Cummins, 2010 #421} {Capic, 2018 #382}.  This suggests that an individual is primarily controlled by how they feel however, there is considerable variation in how an individual adapts to challenges and therefore, SWB levels may change {Lucas, 2007 #535} {Lucas, 2008 #549}. 
 
One explanation of these variations is that of personality. {Costa, 1980 #552}s  theory of personality demonstrated how individuals differ on SWB set-point scores on stable traits of extraversion and neuroticism. Extraverts rated higher on SWB scores than that of introverts and, neuroticism scores were lower than that of emotionally stable individuals. Extraverted individuals tend to create better social experiences which can lead to improved wellbeing {Harris, 2017 #558}.  In contrast, neuroticism can act as a mediator between extraversion and wellbeing leading to lower SWB {Fadda, 2016 #553}. Personality traits may explain why individuals experience life in difference ways and how this influences their wellbeing {Pollock, 2016 #301}. 

However, these examples rely on stable personality traits, challenging life circumstances may destabilise traits and cognitive appraisals may take a salient role. Including a cognitive component may help explain how personality traits differ according to evaluations of satisfaction, and account for underlying mental structures used. One theory to capture these evaluations is that of Multiple discrepancies theory (MDT) {Michalos, 1985 #569} . The hypothesis is that an individual's overall net satisfaction is a function of a perception of ideals between past, present and future desires {Michalos, 2017 #300}. {Jovanovic, 2011 #426} demonstrated that personality doesn’t influence the cognitive component of wellbeing however, neuroticism does influence the affective component. Suggesting that, when personality traits are destabilised the affective component takes control.

Each of these components contribute distinct mechanisms in explaining SWB. Daven tested a model by regressing 5 traits of personality, with affect and cognition on a broad group of adult Australians. {Davern, 2007 #417} demonstrated that the joint contribution of the cognitive-affective model explained 90% of the variance in SWB. Furthermore, personality traits were a weaker contributor than that of cognition and affect; Extraversion and Agreeableness traits were the only significant correlates of SWB. {Blore, 2011 #420} replicated this design with only two personality traits of Neuroticism and Extraversion. The study demonstrated a purely affective model explaining 66% of the variance in SWB. Cognition and the cognitive-affective model proved to be a poor fit to SWB model. {Tomyn, 2011 #418} replicated {Davern, 2007 #417}s design in a group of Australian High school students and found that affect was the better fit then either personality or cognition driven model explaining 80% of the variance. 

While these findings highlight affect as the dominant component, there were some limitations that may have led to unpowered effect sizes. {Blore, 2011 #420} and {Tomyn, 2011 #418} utilised the Ten Item Personality index (TIPI). The TIPI is a short form personality scale designed to perform poorly on reliability factor analysis {Romero, 2012 #503} {Gosling, 2003 #507}. Furthermore, the affect scale used in the {Blore, 2011 #420} was constructed purely for their study and no validity was reported. While {Davern, 2007 #417}s study was comprehensive, no internal validity measurements were reported and the NEO Five Factory  Inventory used in the study has some internal structural issues {Rosellini, 2011 #572}.
 
The present study aimed to retest the relative components of affect, cognition and personality for predicting individual differences in subjective wellbeing. Firstly, by testing the affective, cognitive and personality components on SWB and secondly, to address the limitations from {Blore, 2011 #420} and {Tomyn, 2011 #418}, by utilising a more reliable personality measure of all five personality traits. This study hypothesised that a) mean scores of the PWI will fall within the normative range; b) affect, cognition and personality will predict SWB and; c) affect will have the strongest, unique contribution on predicting SWB.   
  


# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

___Subjective Wellbeing___. Wellbeing was assessed using the Personality Wellbeing Index (PWI) to represent "life as a whole" {Cummins, 2003 #547} {Group, 2013 #567}.  The scale contains seven domains of satisfaction that correspond to represent the overall "How satisfied are you with life as whole?". The domains are: standard of living, health, achieving in life, relationships, safety, community-connectedness, and future security. Participants indicated their level of satisfaction with each domain using a 11-point Likert scale ranging from 0 (No Satisfaction as all) to 10 (Completely Satisfied).  Composite variables were then computed to new variable and labelled as PWI. Cronbachs alpha has been shown to yield small reliability estimates therefore, a more reliable coefficient estimate, omega will be used {McNeish, 2017 #532}{Dunn, 2013 #531}.
Point estimates and confidence intervals for the current sample were $\omega$ = `r round(pwi_reliability$est, digits = 2)`, 95% CI [`r round(pwi_reliability$ci.lower, digits = 2)`, `r round(pwi_reliability$ci.upper, digits = 2)`].  

___Personality___.  Personality was assessed using a 50-item pool from the International Personality Item Pool (IPIP;{Goldberg, 2006 #566} ) which, provides estimates for the Five Factor Model. The five factors and reliability estimates were as follows: Neuroticism $\omega$ = `r round(neo_reliability$est, digits = 2)`, 95% CI [`r round(neo_reliability$ci.lower, digits = 2)`, `r round(neo_reliability$ci.upper, digits = 2)`], Extraversion $\omega$ = `r round(ext_reliability$est, digits = 2)`, 95% CI [`r round(ext_reliability$ci.lower, digits = 2)`, `r round(ext_reliability$ci.upper, digits = 2)`], Openness $\omega$ = `r round(opn_reliability$est, digits = 2)`, 95% CI [`r round(opn_reliability$ci.lower, digits = 2)`, `r round(opn_reliability$ci.upper, digits = 2)`], Agreeableness $\omega$ = `r round(agre_reliability$est, digits = 2)`, 95% CI [`r round(agre_reliability$ci.lower, digits = 2)`, `r round(agre_reliability$ci.upper, digits = 2)`] and Conscientiousness $\omega$ = `r round(conc_reliability$est, digits = 2)`, 95% CI [`r round(conc_reliability$ci.lower, digits = 2)`, `r round(conc_reliability$ci.upper, digits = 2)`]. 

___Affect___. The core construct of SWB is that of Homeostatically Protected Mood based on Russell's (2003)  Core Affect {Cummins, 2010 #421} {Davern, 2007 #417}. Participants were asked to describe their feelings across three areas of life in general: How happy do you feel? How content do you feel? And How alert do you feel? Each item was scored on a Likert rated as 0 (Not at all) to 10 (Extremely). The three items were averaged and a single variable was computed reflecting homeostatically protected mood (HPMood). The reliability of the item for this sample was $\omega$ = `r round(HPmood_reliability$est, digits = 2)`, 95% CI [`r round(HPmood_reliability$ci.lower, digits = 2)`, `r round(HPmood_reliability$ci.upper, digits = 2)`].

___Cognition___. Items from a revised version of Multiple Discrepancies Theory (MDT) was used to evaluate perceived gaps between the current self and various standards of comparison {Michalos, 1985 #569}. This include: What one has and wants (self-wants); what relevant others have (self-other); the best one has had in the past (self-best); what one expected to have 3 years ago (selfprogress); what one expects to have after 5 years (self-future); what one feels they deserve (self-deserves), and what one feels they need (self-needs). Items were averaged across each of the discrepancies to indicate a gap between desired and actual life circumstances. Scores closer to zero indicate less than desired circumstances, scores around five reflect life circumstances close to the current desired level and, scores close to 10 indicate actual circumstances are better than desired. A single MDT variable was computed and the reliability of the items was $\omega$ = `r round(MDT_reliability$est, digits = 2)`, 95% CI [`r round(MDT_reliability$ci.lower, digits = 2)`, `r round(MDT_reliability$ci.upper, digits = 2)`].

# Results

Prior to conducting a hierarchical multiple regression, relevant assumptions of the model were tested. Standardized residuals and qqplots indicated that normality, linearity and multicollinearity were acceptable. However, an examination of Cooks Distance indicated several influential cases which were then removed to prevent extreme leveraging of the model {Cook, 1982 #570}.  

## Descriptives

Means, standard deviations and correlations between variables are presented below in Table 1. The wellbeing average was outside of the normative range of scores (M = 68.80, SD = 10.50) as previous cohort studies shows the average for a sample should fall between 73.43 - 76.43 points {Cummins, 2010 #421}. Neuroticism (r = -.74, p <.01), HPMood (r = .75, p <.01) and MDT (r = .60, p <.01) were all strongly correlated with PWI. Extraversion (r= .40, p <.01), Agreeableness (r = .27, p <.01) and Conscientiousness (r = .29, p <.01) were all moderately correlated with PWI.

```{r, results='asis'}
apa_table(a
          , escape  = FALSE
          , format = "html"
          , caption = "Correlation matrix of the main variables"
          , note    = "Note. M and SD are used to represent mean and standard deviation, respectively. Values in square brackets indicate the 95% confidence interval for each correlation. The confidence interval is a plausible range of population correlations that could have caused the sample correlation (Cumming, 2014). * indicates p < .05. ** indicates p < .01.")
```


# Regression

A three stage hierarchical multiple regression was then conducted with PWI as the dependent variable. HPMood was entered at stage one to control for affect. MDT was entered at stage two and the five personality traits entered as a group at stage three. Regression statistics are presented below with associated change statistics, zero order correlations and regression weights. 

The hierarchical multiple regression revealed that at stage one, HPMood contributed significantly to the model and accounted for 57% of the variation in subjective wellbeing `r {apa_lm_1$full_result$modelfit$r2}`. Introducing MDT at stage two explained 58% of the variance `r {apa_lm_2$full_result$modelfit$r2}` in SWB however, the R2 change was not significant `r {apa_lm_2$statistic$MDT}`. The final step introduced the personality traits: Neuroticism, Extraversion, Openness, Agreeableness and Conscientiousness and explained an additional 10% of the variance in SWB `r {apa_lm_3$full_result$modelfit$r2}`. The final step accounted for 68% of the variance in SWB. In the final adjusted model two out of the seven predictors were statistically significant. Neuroticism had the higher magnitude of effect `r apa_lm_3$full_result$Neuroticism` and the strongest unique contribution to the overall model explaining 8% of the variance in SWB. HPMood `r {apa_lm_3$full_result$HPMood}` provided the next strongest effect size and uniquely explained 4% of the variance in SWB. 

```{r, results='asis'}
apa_table(H_regression$table, added_stub_head = "Variables", note = "Note. A significant b-weight indicates the beta-weight and semi-partial correlation are also significant. b represents unstandardized regression weights. beta indicates the standardized regression weights. sr2 represents the semi-partial correlation squared. r represents the zero-order correlation. LL and UL indicate the lower and upper limits of a confidence interval, respectively.
* indicates p < .05. ** indicates p < .01.
", font_size = "small")
```


Further exploratory model was then conducted excluding non-significant coefficients. A multiple regression analysis showed that HPMood and Neuroticism significantly predicted SWB (`r apa_lm_4$full_result$modelfit$r2`) and accounted for 66% of the variance in SWB.  HPMood proved to have the strongest effect (`r apa_lm_4$full_result$HPMood`) accounting for 12% of the unique variance in SWB followed by Neuroticism (`r apa_lm_4$full_result$Neuroticism`) which contributed 9% of the unique variance.


```{r, results='asis'}
apa_table(Explore_regression$table, added_stub_head = "Variables", note = "This is a note section", font_size = "small")
```




## Discussion

This study aimed to retest the components of affect, cognition and personality on an individuals subjective wellbeing and examine the normative range of wellbeing scores. Despite the position of  {Diener, 2018 #467}  that the cognitive component plays the larger role in wellbeing, previous research has shown that an individuals subjective wellbeing is driven by affect and managed through a neurophysiological process {Cummins, 2010 #421}. Previous testing of these components used unreliable personality inventories and various personality traits were included  {Blore, 2011 #420} {Davern, 2007 #417} {Tomyn, 2011 #418}. Previous results from similar designs either used unreliable personality measures or 

Personality measures have also had varied outcomes on models which led to re-examining 
Which may have influenced other components of the model. 
 
Results from hierarchical regression indicated that affect was the dominant component, consistent with that of {Tomyn, 2011 #418} and  {Blore, 2011 #420}, accounting for 57% of the variation in SWB. The cognition component explained an additional 10% of the variance however, the improvement in the model was not significant. This would explain the significance of when the personality component was entered,  accounting for the additional 10% of the variance in SWB. Indicating that the personality trait of neuroticism takes control of the structure of wellbeing when homeostasis is disrupted {Capic, 2018 #382} {Jovanovic, 2011 #426}.

An unexpected result was that of the samples wellbeing score. The samples mean score (68.80%) was outside the normative range and therefor, evidence there was instability to the homeostatic process {Capic, 2018 #382} {Cummins, 2010 #421}. This becomes evident as the model is processed through after controlling for affect. Based on these results, further exploratory analysis was then conducted with affect and neuroticism as the predictor variables. Both affect and Neuroticism were significant and explained 66% of the variance in SWB. Furthermore, affect contributed the dominant effect size.

A key explanation for these findings is that when SWB is disrupted, the cognitive component drops out and neuroticism becomes the mediating role on wellbeing {Fadda, 2016 #553}. Inconsistent with {Davern, 2007 #417}s hypothesis that when there is a disruption to homeostatic process, cognition will take control and explain more variance in SWB than affect. The importance of this is that personality traits of Neuroticism or Extraversion are likely to play the key role in determining how an individual copes {Costa, 1980 #552} . The components are competing to return the individual to a normal state of functioning {Richardson, 2016 #548}. Therefore, the individual is more likely to rely on emotion rather than cognition when experiencing challenging experiences. 

Under normal conditions, the cognitive-affective model is the preferred structure as demonstrated with {Davern, 2007 #417}. If the sample size is reflective of the adult population, the model should operate under the homeostasis presumption. However, if the sample is likely to be youth or evidence of psychopathology, then a purely affective model would explain their SWB, guided by their level of neuroticism. This has two implications firstly, the cognitive-affective model is useful in guiding public policy, such as broad health interventions and secondly, the affective model has clinical applications. Assessing an individuals wellbeing and personality could highlight how they are likely to cope and therefore, relevant interventions can than be applied.

In light of these findings there were some limitations. Firstly, gender distribution was heavily weighted toward females and may have attenuated the regression line. To date, gender differences in SWB have yet to be fully explored therefore, future designs could implement a hierarchical process by accounting for affect and entering gender separately.  Secondly, both this and previous studies have used a cross sectional methodology. A longitudinal approach may highlight the long term stability of homeostasis address the variation in traits.

Overall, present findings suggest that affect explains the majority of the variance in SWB when wellbeing is in a state of challenge. When the wellbeing is outside the range of normality, trait neuroticism overrides cognition and takes control. In order to address the limitations in this study, future research could explore the relative gender differences in wellbeing and adopt a longitudinal design.

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
