---
title             : "Competing Models of SWB"
shorttitle        : "Models of SWB"

author: 
  - name          : "Aaron L. Willcox"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"

affiliation:
  - id            : "1"
    institution   : "Deakin University"




abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
source("analysis.R")
```
# Introduction
Subjective Wellbeing is the overall evaluation of feeling satisfied absed on cognitive, emotional domains. However, the contributions as to how much of each domain is still under much debate. The Australian Unity Wellbeing Index provides a snapshot of the overwellbeing of Australians each year. This study has a relateively robust score and has deomstrated that, on average, the population PWI average is 75 with a 3.0 percentage points difference. What this suggests is that a persons wellbeing has a set-point as to where their sense of happiness is and therefore can be used to evaluate wellbeing.  


# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

Measures of Employment history were also recorded but were not included in the analyis. 

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results
I need to print and omega symbol $\omega$ and I hope that this worked
maybe if i try this one  or perhaps this one 


# Method
___Subjective Wellbeing___. Wellbeing was assessed using the Personality Wellbeing Index (PWI) to represent "life as a whole" {Cummins, 2003 #547} {Group, 2013 #567}.  The scale contains seven domains of satisfaction that correspond to represent the overall "How satisfied are you with life as whole?". The domains are: standard of living, health, achieving in life, relationships, safety, community-connectedness, and future security. Participants indicated their level of satisfaction with each domain using a 11-point Likert scale ranging from 0 (No Satisfaction as all) to 10 (Completely Satisfied).  Composite variables were then computed to new variable and labelled as PWI. Cronbachs alpha has been shown to yield small reliability estimates therefore, a more reliable coefficient estimate, omega will be used {McNeish, 2017 #532}{Dunn, 2013 #531}.
Point estimates and confidence intervals for the current sample were $\omega$ = `r round(pwi_reliability$est, digits = 2)`, 95% CI [`r round(pwi_reliability$ci.lower, digits = 2)`, `r round(pwi_reliability$ci.upper, digits = 2)`].  

___Personality___.  Personality was assessed using a 50-item pool from the International Personality Item Pool (IPIP;{Goldberg, 2006 #566} ) which, provides estimates for the Five Factor Model. The five factors and reliability estimates were as follows: Neuroticism $\omega$ = `r round(neo_reliability$est, digits = 2)`, 95% CI [`r round(neo_reliability$ci.lower, digits = 2)`, `r round(neo_reliability$ci.upper, digits = 2)`], Extraversion $\omega$ = `r round(ext_reliability$est, digits = 2)`, 95% CI [`r round(ext_reliability$ci.lower, digits = 2)`, `r round(ext_reliability$ci.upper, digits = 2)`], Openness $\omega$ = `r round(opn_reliability$est, digits = 2)`, 95% CI [`r round(opn_reliability$ci.lower, digits = 2)`, `r round(opn_reliability$ci.upper, digits = 2)`], Agreeableness $\omega$ = `r round(agre_reliability$est, digits = 2)`, 95% CI [`r round(agre_reliability$ci.lower, digits = 2)`, `r round(agre_reliability$ci.upper, digits = 2)`] and Conscientiousness $\omega$ = `r round(conc_reliability$est, digits = 2)`, 95% CI [`r round(conc_reliability$ci.lower, digits = 2)`, `r round(conc_reliability$ci.upper, digits = 2)`]. 

___Affect___. The core construct of SWB is that of Homeostatically Protected Mood based on Russell's (2003)  Core Affect {Cummins, 2010 #421} {Davern, 2007 #417}. Participants were asked to describe their feelings across three areas of life in general: How happy do you feel? How content do you feel? And How alert do you feel? Each item was scored on a Likert rated as 0 (Not at all) to 10 (Extremely). The three items were averaged and a single variable was computed reflecting homeostatically protected mood (HPMood). The reliability of the item for this sample was $\omega$ = `r round(HPmood_reliability$est, digits = 2)`, 95% CI [`r round(HPmood_reliability$ci.lower, digits = 2)`, `r round(HPmood_reliability$ci.upper, digits = 2)`].

___Cognition___. Items from a revised version of Multiple Discrepancies Theory (MDT) was used to evaluate perceived gaps between the current self and various standards of comparison {Michalos, 1985 #569}. This include: What one has and wants (self-wants); what relevant others have (self-other); the best one has had in the past (self-best); what one expected to have 3 years ago (selfprogress); what one expects to have after 5 years (self-future); what one feels they deserve (self-deserves), and what one feels they need (self-needs). Items were averaged across each of the discrepancies to indicate a gap between desired and actual life circumstances. Scores closer to zero indicate less than desired circumstances, scores around five reflect life circumstances close to the current desired level and, scores close to 10 indicate actual circumstances are better than desired. A single MDT variable was computed and the reliability of the items was $\omega$ = `r round(MDT_reliability$est, digits = 2)`, 95% CI [`r round(MDT_reliability$ci.lower, digits = 2)`, `r round(MDT_reliability$ci.upper, digits = 2)`].

# Results

Prior to conducting a hierarchical multiple regression, relevant assumptions of the model were tested. Standardized residuals and qqplots indicated that normality, linearity and multicollinearity were acceptable. However, an examination of Cooks Distance indicated several influential cases which were then removed to prevent extreme leveraging of the model {Cook, 1982 #570}.  Means, standard deviations and correlations between variables are presented below in Table 1. The wellbeing average was outside of the normative range of scores (M = 68.80, SD = 10.50) as previous cohort studies shows the average for a sample should fall between 73.43 - 76.43 points {Cummins, 2010 #421}. Neuroticism (r = -.74, p <.01), HPMood (r = .75, p <.01) and MDT (r = .60, p <.01) were all strongly correlated with PWI. Extraversion (r= .40, p <.01), Agreeableness (r = .27, p <.01) and Conscientiousness (r = .29, p <.01) were all moderately correlated with PWI. 

A three stage hierarchical multiple regression was then conducted with PWI as the dependent variable. HPMood was entered at stage one to control for affect. MDT was entered at stage two and the five personality traits entered as a group at stage three. Regression statistics are presented below with associated change statistics, zero order correlations and regression weights. 

The hierarchical multiple regression revealed that at stage one, HPMood contributed significantly to the model,  and accounted for 57% of the variation in subjective wellbeing `r apa_lm_1$full_result$modelfit`  . Introducing MDT at stage two explained 58% of the variance `r apa_lm_2$full_result`  in SWB however, the R2 change was not significant`r apa_lm_2$statistic$MDT`. The final step introduced the personality traits: Neuroticism, Extraversion, Openness, Agreeableness and Conscientiousness and explained an additional 10% of the variance in SWB`r apa_lm_3$full_result$modelfit` . The final step accounted for 66% of the variance in SWB. In the final adjusted model two out of the seven predictors were statistically significant. Neuroticism had the higher magnitude of effect `r apa_lm_3$full_result$HPMood` and the strongest unique contribution to the overall model explaining 8% of the variance in SWB. HPMood  provided the next strongest effect size and uniquely explained 4% of the variance in SWB.


```{r, results='asis'}
apa_table(H_regression$table, added_stub_head = "Variables", note = "This is a note section", font_size = "small")
```

Then on the next page we have here a loverly correlation table for me to check it and see if the covaiates match with the regression model
```{r, results='asis'}
apa_table(a
          , escape  = FALSE
          , format = "html"
          , caption = "Correlation matrix of the main variables"
          , note    = "Note. M and SD are used to represent mean and standard deviation, respectively. Values in square brackets indicate the 95% confidence interval for each correlation. The confidence interval is a plausible range of population correlations that could have caused the sample correlation (Cumming, 2014). * indicates p < .05. ** indicates p < .01.")
```



# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
